---
title: "Simple Markdown Example"
subtitle: "PDF/LaTex version"
author: 
  - "Jenny Rieck"
  - "Derek Beaton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    keep_tex: true
urlcolor: blue
---

```{r libraries, echo=FALSE, message=FALSE}

require(kableExtra)
require(gridExtra)
require(tidyverse)
require(ExPosition)
require(GSVD)
require(knitr)
require(reticulate)

if(!require(covstatis)){
  install_github("jennyrieck/C-MARINeR", subdir = "/code/covstatis")
  library(covstatis)
}

```

```{r rmarkdown_setup, echo=FALSE, message=FALSE}

opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

```

```{r source_functions, echo=FALSE, message=FALSE}



```

```{r load_data, echo=FALSE, message=FALSE}

load(file=paste0(Sys.getenv("ADNI_FOLDER"),"\\","amerge_subset.rda"))
load(file=paste0(Sys.getenv("ADNI_FOLDER"),"\\","variable_type_map.rda"))

```

# Introduction

For reference, see [RMarkdown Driven Development](https://emilyriederer.netlify.com/post/rmarkdown-driven-development/) by Emily Riederer for a comprehensive overview of how to best structure RMarkdown for projects, packages, and other development-driven tasks. 

In general for this RMarkdown file we follow Emily's fourth example with a directory structure, minimized redundancies, and heavy-duty code elsewhere. This RMarkdown, generally, serves as place to describe, analyze, and visualize our data. Thus, this text is shown at the top of the output, but actually appears after several `R` code chunks, which exist between the `Introduction` heading and YAML header information (which you see as title, authors, dates, etc...).

Most of the RMarkdown files in this directory will show generally the same content, but help highlight the different ways you can use RMarkdown, `knitr`, `pandoc`, `LateX`, and various package built for those, such as `beamer` (`LateX`) for presentations, and [`papaja`](https://github.com/crsh/papaja) & [`rticles`](https://github.com/rstudio/rticles) (R/RMarkdown) for writing manuscripts that export to `LaTeX` or MS Word. If you decide to write MS Word documents through RMarkdown, you should also use the [`redoc` package](https://github.com/noamross/redoc). 

## R chunks: A word of caution

It is good practice to name your `R` chunks. If you do not, then the `R` chunks will still produce the intended material. However, when you do name them, you should ensure they have unique names (else, you will likely see some cryptic and not always informative error messages).


# Tables

There are multiple approaches and packages to help visualize tables or tabular information. Let's start by looking at a simple summary of all the continuous variables. First, we will visualize the summary table through two methods within R: `knitr::kable` and the `kablExtra` package, followed by `grid` and `gridExtra`. Next, we will use the same data and illustrate what happens when we pass it to Python through `reticulate`. 

In this section we wil also show the code chunks that generate these tables and visuals, which are embedded in the RMarkdown document. 

## knitr and kableExtra

To make HTML and LaTeX tables in RMarkdown, one of the easiest and most common options is through `knitr`. The `knitr` package is, effectively, the tool to make RMarkdown documents go from R & RMarkdown (plus other code and LaTeX) into PDFs or HTML pages. We'll start with the `knitr::kable()`.


```{r knitr_table_setup}

example_table <- amerge_subset[,c("AGE","MOCA","CDRSB","WholeBrain","Hippocampus","MidTemp")]
example_table_Dx <- amerge_subset[,c("DX","AGE","MOCA","CDRSB","WholeBrain","Hippocampus","MidTemp")]

```

```{r knitr_kable_example}

kable(summary(example_table))

```

But that is not particularly nice looking. So we can use some parameters to make this table look better (which depend on having LaTeX).


```{r knitr_kable_nicer_example}

kable(summary(example_table), format="latex", booktabs=T) 

```

With `booktabs` and `latex` format, we've made the table look a little better. But can we make it look even better than that? We can with `kableExtra`. 

```{r knitr_kableExtra_example}

kable(summary(example_table), format="latex", booktabs=T) %>% kable_styling(font_size=10, position = "center")

```

We can take the table look even further with additional options, like "stripes".

```{r knitr_kableExtra_better_example}

kable(summary(example_table), format="latex", booktabs=T) %>% kable_styling(font_size=10, position = "center", latex_options = "striped")

```

Given that we have redundant information in the table (min/max, etc...) we can do a better job and make an even nicer table with an `apply()`, and then use multiple `kable` and `kableExtra` features to make a really nice table.

```{r kniter_even_better_example}

better_example_table <- apply(example_table, 2, summary)

kable(better_example_table, format="latex", booktabs=T, digits=2) %>% 
  kableExtra::add_header_above(c("Statistic" = 1, "Demographic" = 1, "Clinical" = 2, "Brain" = 3)) %>%
  kable_styling(font_size=10, position = "center", latex_options = "striped") %>%
  row_spec(0, angle = 15, bold=T)

```


## grid and gridExtra

Sometimes we need tables to be graphics. This is where the `grid` and `gridExtra` packages come in. The `grid` package uses `grob`s to turn items---tables, figures, all sorts of things---into configurable parts of a figure.

```{r grid_and_gridExtra}

grid.table(better_example_table)

```

But it's clear we have a few extra things we should do to make this figure of a table look better.

```{r grid_and_gridExtra_round}

better_example_table_round <- apply(better_example_table, 2, format, digits=2, nsmall=2)

grid.table(better_example_table_round)

```

However, the `grid` and `gridExtra` packages can be difficult to customize many of the pieces. Therefore it might be easier to stick to the `LaTeX` approaches with `kable` or it is well worth checking out the [`gt` package](https://github.com/rstudio/gt).


## Python via reticulate

What if you now love RMarkdown but are still really into Python? Not a problem. The [`reticulate` package](https://rstudio.github.io/reticulate/) has you covered. It's an R package to connect to your `Python` installation and bring the data or results back into `R`, but with some of the same features as you're used to in either `Python` or `R`. Let's start out with a *head to head* of `R`'s `head()` vs. `Python`'s `.head()`.

In `R` we write the call in RMarkdown as if we would normally in `R`:

```{r r_head_example}
head(example_table)
```

Unlike the previous code chunks, we have to tell RMarkdown that the language it should expect is `python` instead of `R`. 

```{python r_python_example}
print(r.example_table.head())
```

For `Python` via `R` we also need to use the `.` (dot notation) because of `Python`'s object oriented approach.  From the `r` object, we get the `example_table` attribute and then perform the `head` method. That's because `Python` needs to know about the object coming from `R`. 


### Here be .dragons

For those more familiar with `R` style that stems from the [`S` language origin]() or [Google's style guide](), the `.` can be a substantial source of confusion both here and in the `tidyverse`. In base `R`, it is a valid character for user defined items. But in base `R` it is *also* used for objects and classes (see, e.g., `.print()`). In the `tidyverse` the `.` has a special purpose when alone (not amongst other characters), often as a placeholder for where to pass a variable as an argument into a function. 

In `R`: use `.` either with caution or reckless abandon.


### Passsssing back and foRth

That subtitle is a stretch! With the `reticulate` package and `R` we can pass items between the the two languages/environments. And, depending on which language, we use that language's preferred/standard approach of referring to attributes or objects. In this next example we call into the `.describe()` method in `Python`, which is similar to `R`'s `summary()`, but allows/requires us to define certain parameters. 

```{python describe_example}
perc = [.25, .50, .75] 
desc = r.example_table.describe(percentiles = perc)

print(desc)

```

We can visualize the results directly using the `print()` method. But like previous results shown in `R`, we can do a lot better. We can pass the `desc` object back to `R` with the `$` notation (generally for lists or simple classes in `R`). When we retrieve an object from python (via `py$`), we can then do what we would usually do in `R`. Here, we go back to `kable()` and `kable_styling()` with some of the parameters to make very nice looking `LaTeX` tables.


```{r python_describe_via_kable}
kable(py$desc, digits = 2, format="latex", booktabs=T) %>% kable_styling(font_size=10, position = "center", latex_options = "striped")
```


# Analyses


# Visualizations & Graphics


